\chapter*{Preface}

\section*{About the Book}
This book is sort of a big notebook to make (or force) the author to self study and understand the field of Probability and Statistics.
Concepts and topics are explained with details and examples. Almost all theorems and lemmas have proofs. There are some exceptions on basic or similiar theorems where proof is only a sketch.
I have to note that this book is an educational and fun project for the author himself. Through the book, author tries to explain the topics to \textit{himself} .Be careful using the book as the main learning material, since the writer himself is not an expert in the field, there may be mathematical errors in the book.
\par

I like to explain the mathematical concepts in more ``traditional'' way. I don't like long and complex theorems, lemmas with comically big proofs that reader must pray to understand. Through the book, I try to explain the concepts in everyday language. Of course, rigorous proofs are also provided as they are still an important part of mathematics.
\par
To learn the field and write this book, I used various books from known authors, countless mathematics forums about statistics and probability, and wikipedia (duh-duh) articles. These are some of the books I used majority of time:

\begin{itemize}
    \item Larry Wasserman - All of Statistics - A Concise Course in Statistical Inference.
    \item Dimitri Bertsekas And John N Tsitsiklis - Introduction To Probability
    \item Mathematical Statistics with Applications by Dennis Wackerly, William Mendenhall, and Richard L. Scheaffer
    \item Joseph K. Blitzstein - Introduction to Probability
    \item Ross, Sheldon - First Course in Probability
  \end{itemize}
I want to note that I did not, by any means, plagiarize any contents, diagrams or other things. I simply wrote whatever I learnt through the the brainstorm I had. Theorems and proofs may be similar, but I believe it is acceptable since I can't rigorously find another way of defining theorems and proving them.
  

\subsubsection*{Book's source}
Maybe you may already know this, this book is fully open source with its pictures and tex file shared in author's \href{https://github.com/JosephMehdiyev}{github}. You may use the source code for whatever purposes you want to use it for. If you want to contribute, please send a pull request from the github.Currently the book is in development.

\section*{How to use the Book}
As the book is precise and short, you may use the book as a revisit or a secondary material. The book shortly and simply explains the concepts and ideas. Important concepts' proofs are provided. However, other proofs explaned in sentences rather than other classic rigorous proofs.

\section*{Coding stuff}
The statistical images are being generated by \textbf{python's matplotlib}, while other sort of diagrams are mix of \textbf{latex's tikz} or $\textbf{matplotlib}$. Moreover, there are practical examples  with $\textbf{python}$ of probability and statistical concepts through the book. You can get more information from the book's github page.
